Concurrent Programming in Harmony

### 2021-03-21 book/paper.tex
Latest

In \texttt{release\_rlock}, notice that \texttt{notify}(?\textit{w\_cond})
is invoked when there are no readers left, \emph{without} checking
if there are writers waiting to enter.  This is
ok, because calling \texttt{notify} is a no-op if no thread is
waiting.

\texttt{release\_wlock} executes \texttt{notifyAll}(?\textit{r\_cond})
as well as \texttt{notify}(?\textit{w\_cond)}.  Because we
do not keep track of the number of waiting readers or writers, we
have to conservatively assume that all waiting readers can enter,
or, alternatively, up to one waiting writer can enter.  So
\texttt{release\_wlock} wakes up all potential candidates.  There
are two things to note here.  First, unlike split binary semaphores
or Hoare monitors, where multiple waiting readers would have to be
signaled one at a time in a baton-passing fashion (see
\autoref{fig:RWsplitsema}), with Mesa monitors all readers are
awakened in one fell swoop using \texttt{notifyAll}.  Second, both
readers and writers are awakened---this is ok because both execute
\texttt{wait} within a \textbf{while} loop, re-checking the condition
that they are waiting for.  So if both type of threads are waiting,
either all the readers get to enter next or one of the writers gets
to enter next.  (If you want to prevent waking up both readers and
a writer, then you can keep track of how many threads are waiting
in the code.)

@@

The following people contributed by making comments on or finding bugs in
early drafts of the book:
Alex Chang,
Anneke van Renesse,
Brendon Nguyen,
Hartek Sabharwal,
Heather Zheng,
Jacob Brugh,
Lorenzo Alvisi,
Saleh Hassen,
Sunwook Kim,
Terryn Jung,
Trishita Tiwari,
William Ma,
Xiangyu Zhang,
Yidan Wang,
Zhuoyu Xu,
and
Zoltan Csaki.

>>>>

In \texttt{read\_release}, notice that \texttt{notify}(?\textit{w\_cond})
is invoked when there are no readers left, \emph{without} checking
if there are writers waiting to enter.  This is
ok, because calling \texttt{notify} is a no-op if no thread is
waiting.

\texttt{write\_release} executes \texttt{notifyAll}(?\textit{r\_cond})
as well as \texttt{notify}(?\textit{w\_cond)}.  Because we
do not keep track of the number of waiting readers or writers, we
have to conservatively assume that all waiting readers can enter,
or, alternatively, up to one waiting writer can enter.  So
\texttt{write\_release} wakes up all potential candidates.  There
are two things to note here.  First, unlike split binary semaphores
or Hoare monitors, where multiple waiting readers would have to be
signaled one at a time in a baton-passing fashion (see
\autoref{fig:RWsplitsema}), with Mesa monitors all readers are
awakened in one fell swoop using \texttt{notifyAll}.  Second, both
readers and writers are awakened---this is ok because both execute
\texttt{wait} within a \textbf{while} loop, re-checking the condition
that they are waiting for.  So if both type of threads are waiting,
either all the readers get to enter next or one of the writers gets
to enter next.  (If you want to prevent waking up both readers and
a writer, then you can keep track of how many threads are waiting
in the code.)

@@

The following people contributed by making comments on or finding bugs in
early drafts of the book:
Alex Chang,
Anneke van Renesse,
Brendon Nguyen,
Hartek Sabharwal,
Heather Zheng,
Jacob Brugh,
Lorenzo Alvisi,
Maria Martucci,
Saleh Hassen,
Sunwook Kim,
Terryn Jung,
Trishita Tiwari,
William Ma,
Xiangyu Zhang,
Yidan Wang,
Zhuoyu Xu,
and
Zoltan Csaki.
